---
title: "p8105_hw2_rt2712"
author: "Rachel Tsong"
date: "5 October 2018"
output: github_document
---

# Problem 1

## Step 1

Load packages
```{r}
library(tidyverse)
```


## Step 2

Load data, clean names, select necessary columns

```{r}
subway_data = (
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select(line, station_name, station_latitude, station_longitude, entrance_type, entry, route1:route11, vending, ada) %>%
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE))
)
```

### Summary of dataset

First I imported the data from the .csv file and cleaned the names so that the dataset is easier to work with. After that, I selected the columns for line, station name, station latitude and longitude, entrance type, routes served, vending, and ADA compliance. Lastly, I changed the entry column into a logical vector. The resulting data set is 1868 x 19. Right now, the data frame is not yet tidy because there are variables in the column names: route1-route11. These columns need to be reformatted so that each variable has its own column, each observation has its own row, and there is a single value in every cell.

## Step 3 

### Answer questions:

**How many distinct stations?**
```{r}
distinct(subway_data, line, station_name)
```

There are 465 unique stations. 

**How many stations are ada compliant?**
```{r}
filter(
  (distinct(subway_data, line, station_name, .keep_all = TRUE)),
  ada == TRUE)
```

84 distinct stations are ADA compliant.

**What proportion of entrances/exits without vending allow entrance?**
```{r}
filter(subway_data, vending == "NO", entry == TRUE)
```

69 entrances/exits without vending allow entrance.

## Step 4

Reformat route # and route name, answer questions
```{r}
subway_data_tidy= gather(subway_data, key = route_number, value = route, route1:route11)

filter(
  (distinct(subway_data_tidy, line, station_name, .keep_all = TRUE)),
  route == "A"
)

filter(
  (distinct(subway_data_tidy, line, station_name, .keep_all = TRUE)),
  route == "A", ada == TRUE
)
```

There are 60 distinct subway stations that serve the A train, and 17 of these stations that serve the A train are ADA compliant.






































