---
title: "p8105_hw2_rt2712"
author: "Rachel Tsong"
date: "5 October 2018"
output: github_document
---

# Problem 1

## Step 1

Load packages
```{r}
library(tidyverse)
```


## Step 2

Load data, clean names, select necessary columns

```{r}
subway_data = (
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select(line:entry, vending, ada) %>%
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE))
)
```

### Summary of dataset

First I imported the data from the .csv file and cleaned the names so that the dataset is easier to work with. After that, I selected the columns for line, station name, station latitude and longitude, entrance type, routes served, vending, and ADA compliance. Lastly, I changed the entry column into a logical vector. The resulting data set is 1868 x 19. Right now, the data frame is not yet tidy because there are variables in the column names: route1-route11. These columns need to be reformatted so that each variable has its own column, each observation has its own row, and there is a single value in every cell.

## Step 3 

### Answer questions:

**How many distinct stations?**
```{r}
distinct(subway_data, line, station_name)
```

There are 465 unique stations. 

**How many stations are ada compliant?**
```{r}
filter(
  (distinct(subway_data, line, station_name, .keep_all = TRUE)),
  ada == TRUE)
```

84 distinct stations are ADA compliant.

**What proportion of entrances/exits without vending allow entrance?**
```{r}
filter(subway_data, vending == "NO")

filter(subway_data, vending == "NO", entry == TRUE)
```

I ran the first line of the code chunk above to tell me how many stations don't have vending and the second line to tell me how many stations that do not offer vending allow entrance. I can conclude from the resulting tibbles that 69/183 or 37.7% of stations that do not offer vending do allow entrance.

## Step 4

Reformat route # and route name, answer questions
```{r}
subway_data_tidy = gather(subway_data, key = route_number, value = route, route1:route11)

filter(
  (distinct(subway_data_tidy, line, station_name, .keep_all = TRUE)),
  route == "A"
)

filter(
  (distinct(subway_data_tidy, line, station_name, .keep_all = TRUE)),
  route == "A", ada == TRUE
)
```

There are 60 distinct subway stations that serve the A train, and 17 of these stations that serve the A train are ADA compliant.

# Problem 2

```{r read and clean sheet 1}
library(readxl)

trash_wheel = (
  read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = 1, range = "A2:N256") %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>% 
  mutate(sports_balls = as.integer(round(sports_balls)))
)
```

```{r read and clean precip data}
precip_2017 = (
  read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = 3, range = "A2:B14") %>% 
  janitor::clean_names()
)

precip_2016 = (
  read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = 4, range = "A2:B14") %>% 
  janitor::clean_names() 
)



```


































